# Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

-   Answers to the research questions
-   Different methods considered
-   Competing approaches
-   Justifications

## Lait Equitable Analysis

### Pareto Principle

The Pareto Principle, often known as the 80/20 rule, asserts that a small proportion of causes, inputs, or efforts usually lead to a majority of the results, outputs, or rewards. Applied to a business context where approximately 20% of the sales account for 80% of the revenues, this principle can help in identifying and focusing on the most profitable aspects of a business.

Evidence from Research:

Sales and Customer Concentration: Research has consistently shown that a significant portion of sales often comes from a minority of customers or products. For instance, an analysis across 22 different consumer packaged goods categories found an average Pareto ratio (PR) of .73, indicating that the top proportion of products/customers often account for a disproportionately high share of sales or profits 
Source - [Kim, Singh, & Winer, 2017](https://doi.org/10.1007/S11002-017-9442-5)

Decision Making and Resource Allocation: The Pareto Principle helps in decision-making by highlighting areas where the greatest impact can be achieved. For example, focusing on the top-performing products or customers can optimize resource allocation and maximize profits 
Source - [Ivančić, 2014](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2841896)

Market and Profit Concentration: Another study noted that a small number of customers are often responsible for a large portion of sales, which supports the strategic focus on these customers to boost profitability and efficiency 
Source- [McCarthy & Winer, 2018](https://doi.org/10.2139/ssrn.3264425)

Conclusion:
Applying the Pareto Principle in a business context where a minority of sales drives the majority of revenue can lead to more focused and effective business strategies, optimizing efforts towards the most profitable segments. This approach not only simplifies decision-making but also enhances resource allocation, ultimately leading to increased profitability.

#### Steps

1.  Calculating the total sales across all locations for both 2022 and 2023.
2.  Ranking locations by sales to see the cumulative contribution of each location towards the total.
3.  Identifying the point where approximately 20% of the locations contribute to around 80% of the sales.

```{python}
# Calculate the total sales for each year and the combined total to apply Pareto Principle
merged_sales_data['Combined Sales'] = merged_sales_data['Total Sales 2022'] + merged_sales_data['Total Sales 2023']

# Sort locations by combined sales
pareto_data = merged_sales_data.sort_values(by='Combined Sales', ascending=False)

# Calculate cumulative sales
pareto_data['Cumulative Sales'] = pareto_data['Combined Sales'].cumsum()

# Calculate the total of combined sales
total_combined_sales = pareto_data['Combined Sales'].sum()

# Calculate the percentage of cumulative sales
pareto_data['Cumulative Percentage'] = 100 * pareto_data['Cumulative Sales'] / total_combined_sales

# Find the point where about 20% of the locations contribute to approximately 80% of the sales
pareto_data['Location Count'] = range(1, len(pareto_data) + 1)
pareto_data['Location Percentage'] = 100 * pareto_data['Location Count'] / len(pareto_data)

# Plotting the Pareto curve
plt.figure(figsize=(12, 8))
cumulative_line = plt.plot(pareto_data['Location Percentage'], pareto_data['Cumulative Percentage'], label='Cumulative Percentage of Sales', color='b', marker='o')
plt.axhline(80.2, color='r', linestyle='dashed', linewidth=1)
plt.axvline(33.3, color='green', linestyle='dashed', linewidth=1)
plt.title('Pareto Analysis of Sales Across Locations')
plt.xlabel('Cumulative Percentage of Locations')
plt.ylabel('Cumulative Percentage of Sales')
plt.legend()
plt.grid(True)
plt.show()
```
Given this graph 33.2% of Manor locations are contributing to 80% of sales. This deviates from the typical Pareto 80/20 distribution, but it still shows a concentration of sales among a subset of stores.

#### Observations

We will identify the top 33.3% of locations based on their cumulative sales contribution. This means selecting the smallest number of locations that together account for at least 80% of the total sales.

The top-performing 33.3% of Manor locations that contribute to the majority of sales are:
```{python}
# Calculate the threshold for the top 33.3% of locations
top_third_index = int(len(pareto_data) * 0.34)

# Identifying the top 33.3% of stores contributing to at least 80% of sales
top_performing_stores = pareto_data.head(top_third_index)
top_performing_stores
```

### Understanding Success Factors of Top-Performing Stores

#### Correlating Political Parties with Milk Sales
```{r}
# Calculate correlation coefficients for each party
correlation_df <- data.frame(Party = c("PLR", "PS", "UDC", "Centre", "Verts"),
                             Correlation = sapply(merged_data[, 4:8], function(x) cor(x, merged_data$`2023`)))

# Print the correlation coefficients
print(correlation_df)

# Create a matrix of plots for each party
party_plots <- lapply(names(merged_data)[4:8], function(party) {
  ggplot(merged_data, aes_string(x = "`2023`", y = party)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue") +
    labs(x = "Annual Sales", y = paste(party, "Party Presence (%)"), title = paste("Correlation:", party, "Party vs. Sales")) +
    theme_minimal()
})

# Arrange the plots in a matrix layout
matrix_plot <- gridExtra::grid.arrange(grobs = party_plots, ncol = 2)
matrix_plot
```

